Self-tunable approximated explicit MPC: Heat exchanger implementation and analysys

### 2023-01-27 paper/elsarticle/main.tex
Review: Sec.4

In this section, the results of the proposed tuning method are presented on an experimental implementation. The tuning strategy utilizes tuning parameter based n the size of reference change (Section~\ref{sec:self_tunable_delta_ref}) and the scaling of tuning parameter based on splitting the interval of the parameter and assigning the interval parts to specific operating conditions (Section~\ref{sec:self_tunable_rho_scaling}).

The plant on which the control was implemented is a laboratory liquid-to-liquid plate heat exchanger Armfield Process Plant Trainer PCT23~\cite{pct23}, see Fig.~\ref{fig:HE}. The cold feed as well as heating medium are transported to the heat exchanger by two peristaltic pumps. The flow rate of the feed is constant, while the aim is to track the reference value of its temperature. Therefore, the controlled output is the feed temperature $T$ at the outlet of the heat exchanger and the control input is the voltage $U$ corresponding to the power of the heating medium pump. The voltage is set in percentage, while it is constrained from 20\% to 100\%. As the heat exchange is a nonlinear and asymmetric process~\cite{Liptak}, this heat exchanger represents a suitable candidate for the presented controller tuning strategy.

The matrices of the linear state-space model of the plant, discretized with sampling time $T_\mathrm{s}$ = 1\,s, are
\begin{subequations}
	\label{eq:model_A_B} 
	\begin{eqnarray}
		A&=&\begin{bmatrix}
			0.839
		\end{bmatrix}, \\
		B&=&\begin{bmatrix}
			0.039
		\end{bmatrix}, \\
		C&=&\begin{bmatrix}
			1
		\end{bmatrix}, \\
		D&=&\begin{bmatrix}
			0
		\end{bmatrix}.
		\end{eqnarray}
\end{subequations}

The constraints are considered in the terms of control inputs, i.e.
\begin{eqnarray}
\label{eq:u_const}
	-15\% \le u \le 65\%.
\end{eqnarray}

Note, the variable $u$ represents the control inputs in the deviation form. The values of feed temperature and voltage of the heating medium pump corresponding to zero steady state are respectively $T^\mathrm{s}$~=~35 $^{\circ}\mathrm{C}$ and $U^\mathrm{s}$~=~35\%.

The penalty matrices of the problem Eq.~\eqref{eq:mpc_problem} were systematically tuned and the corresponding control was implemented on the laboratory heat exchanger for every MPC setup. The aim was to determine, which penalty matrix will be the tuned one. Based on observations, the most significant effect on the control trajectories had tuning the penalty matrix $Q_\mathrm{y}$, while still preserving a satisfactory control performance, i.e., without steady-state control error and significant oscillations around the setpoint. The boundary values of matrix $Q_\mathrm{y}$ were tuned as $Q_\mathrm{y, L}$ = 100 and $Q_\mathrm{y, U}$ = 1000. The built-in integrator was penalized with $Q_\mathrm{x}$ = 1 and the control input with $R$ = 10. The prediction horizon $N$ was 20 steps long. The explicit model predictive controllers were constructed in MATLAB R2020b using the Multi-Parametric Toolbox 3~\cite{mpt_conf}. 
The controllers were implemented to track a time-varying reference. For the first 200 seconds, the system was in the steady state. After that, the reference changed its value twice upwards and twice downwards. The reference changes also acquired different sizes in order to examine the proposed tuning method as it is dependent on the size of the reference step change. The control results for both control setups are compared in Fig.~\ref{fig:CV_boundaries} for the controlled outputs, and in Fig.~\ref{fig:MV_boundaries} for the control inputs.

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{images/CV_boundaries}
		\caption{Controlled output trajectory for two boundary controllers. The solid lines represent the controlled temperature and the dashed line represents the reference value.}
		\label{fig:CV_boundaries}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{images/MV_boundaries}
		\caption{Control input trajectory for two boundary controllers. The solid lines represent the voltage and the dashed lines represent the constraints.}
		\label{fig:MV_boundaries}
	\end{center}
\end{figure}

The trajectories in Fig.~\ref{fig:CV_boundaries} show the asymmetric nature of controlling the process of heat exchange mainly when observing the overshoots and undershoots. When applying the control associated with the lower bound $Q_\mathrm{y, L}$, significant undershoots are present when tracking the reference downwards, i.e., when the reference change is negative. On the contrary, when implementing the controller associated with $Q_\mathrm{y, U}$, the undershoots are negligible, but significant overshoots can be seen when tracking the reference upwards.

These observations form the base for the strategy of self-tuning the penalty matrix $Q_\mathrm{y}$. The strategy follows the ideas summarized in Section~\ref{sec:self_tunable_new}. Utilizing the boundary controller with the penalty matrix $Q_\mathrm{y,L}$ is preferred when the reference changes upwards. On the contrary, the controller associated with $Q_\mathrm{y,U}$ is preferred in control with negative reference changes. The splitting value of the tuning parameter was chosen in the middle of the interval, i.e., $\rho_{\mathrm{s}} = 0.5$. The last parameter that needed to be set was the maximal possible size of the reference step change $d_{\max}$, which was set to $15^{\circ}\mathrm{C}$.

The control results of the selt-tunable technique compared to the boundary controllers can be seen in Fig.~\ref{fig:CV} for the controlled outputs, and in Fig.~\ref{fig:MV} for the control inputs. It can be seen that the tuned controller combined the benefits of the two boundary controllers. The overshoots and undershoots were reduced, as in the first half of control the penalty matrix $Q_\mathrm{y}$ acquired value from the first half of the penalty interval. When tracking the reference with negative change, the penalty matrix acquired the values from the second half of the interval, i.e., closer to the upper bound $Q_\mathrm{y,U}$. 
The similarity with the boundary controllers can be seen also on the control inputs trajectory. Note, the constraints on the input variable were satisfied as they were scaled using linear interpolation based on the boundary controllers which are constructed considering the contraints.

TODO: what about state constraints?

@@

The control performance was also investigated quantitatively. The Table~\ref{tab:control_performance} summarizes the evaluated control performance criteria for the tuned and two boundary controllers. The control performance is evaluated for every reference step change respectively. The criteria are integral square error $ISE$, maximal overshoot/undershoot $\sigma_{\mathrm{max}}$, settling time $t_{\epsilon}$ for 5\%-neighbourhood of the reference value, and the volume of the heating medium utilized for the corresponding control. To provide better readability of the results, the best (i.e. minimal) values are written in bold.

\begin{table}[h!]
	\begin{center}
		\caption{Control performance criteria.}
		\label{tab:control_performance}
		\begin{tabular}{c|c|c|c|c|c} 
			Reference step change & $Q_\mathrm{y}$ & $ISE$ [$^{\circ}\mathrm{C}^2$\,s] & $\sigma_{\mathrm{max}}$\,[\%] & $t_{\epsilon}$\,[s] & $V$\,[l] \\
			\hline
			\multirow{2}{*}{1} & 1000 & 714 & 33.50 & 16.5 & \textbf{2.12} \\
			    & 100 & 867 & 16.65 & 12.5 & 2.36 \\ 
			    & tuned & \textbf{678} & \textbf{15.19} & \textbf{9.5} & 2.38 \\ 
			\hline
			\multirow{2}{*}{2} & 1000 & 365 & 47.20 & \textbf{5} & \textbf{2.49} \\
			    & 100 & 606 & 23.25 & 26.5 & 3.19 \\ 
			    & tuned & \textbf{248} & \textbf{19.13} & 9.5 & 3.35 \\ 
			\hline
		    \multirow{2}{*}{3} & 1000 & 245 & \textbf{18.92} & \textbf{6.5} & \textbf{2.00} \\
				& 100 & 398 & 79.64 & 31 & \textbf{2.00} \\ 
				& tuned & \textbf{186} & 24.59 & \textbf{6.5} & 2.10 \\ 
			\hline
			\multirow{2}{*}{4} & 1000 & 1024 & 18.43 & 22.5 & 0.94 \\
				& 100 & 1402 & 41.87 & 90 & \textbf{0.93} \\ 
				& tuned & \textbf{967} & \textbf{16.49} & \textbf{18.5} & 1.10  
		\end{tabular}
	\end{center}
\end{table}

It can be seen in Table~\ref{tab:control_performance}, that real-time tuning of the controller helped to improve two to three criteria when tracking each reference value. The cost for this is the energy associated with control. Although the integral square error, maximal overshoot/undershoot and settling time decreased, the volume of consumed heating medium did not. Nevertheless, the average deterioration in the terms of consumed heating medium is approximately~17\%.

The improvement or deterioration in percentage of using the tunable controller relative to the second best setup is summarized in Table~\ref{tab:improvement} for every reference step change separately. The negative numbers represent deterioration compared to the best controller setup in the corresponding reference tracking.

\begin{table}[h!]
	\begin{center}
		\caption{Improvement or deterioration of the control performance in percentage.}
		\label{tab:improvement}
		\begin{tabular}{c|c|c|c|c} 
			Reference step change & $ISE$ & $\sigma_{\mathrm{max}}$ & $t_{\epsilon}$ & $V$ \\
			\hline
			1 & 5.04 & 8.77 & 24.00 & -12.26 \\ 
			\hline
			2 & 32.05 & 17.72 & -90.00 & -34.54 \\ 
			\hline
			3 & 24.08 & -29.97 & 0 & -4.48 \\ 
			\hline
			4 & 5.57 & 10.53 & 17.78 & -18.28  
		\end{tabular}
	\end{center}
\end{table}

It can be seen that implementing a tunable controller leads to improved control performance in many criteria. Utilizing a controller with a scalable aggressivity according to the operating conditions leads in general to higher accuracy, lower overshoot and faster achieving the reference value. %It should be also noted that the deterioration of every criterium is calculated subject to the best out of \textit{two} other controllers. It follows that the evaluation is strict as in practice only one controller would be available.
Although the volume of the heating medium is not saved in this control scenario, the improved control performance may lead to other benefits.

@@

\section{Conclusion}
\label{sec:conclusion}
This paper deals with the experimental implementation of tunable approximated explicit model predictive control and provides a technique for self-tuning controller design. Based on the current reference value, the tuning parameter is scaled using linear interpolation. The previously published paper related to the self-tunable explicit MPC suggested tuning based on the distance of the reference value from the steady state. This paper presents a novel idea of tuning based on the size of reference step change. The tuning procedures aim to compensate for the nonlinear behavior of the controlled system. The tuning parameter is updated when the reference changes. It is calculated as the ratio between the size of the reference change and the maximal possible size of the reference change, which is specified before operation. Another novel contribution comes with splitting the interval of tuning parameter into two ranges, while both ranges are assigned to different operating conditions. The proposed method is implemented on a laboratory heat exchanger with nonlinear and asymmetric behavior. The non-symmetry makes the plant a suitable candidate for splitting the interval of the tuning parameter. The decision criterium is negativity or positivity of reference change. When the reference changed upwards, the control input was tuned in the first part of the interval and approached the boundary controller associated with the lower bound on the penalty matrix. On the contrary, when the reference changed downwards, the control input was tuned to approach the control input from the boundary controller with the upper bound on the penalty matrix. To investigate the control results properly, the control performance was evaluated quantitatively. Compared to control with only one controller, the self-tunable technique helped to decrease the maximal overshoots/undershoots, integral square error, and settling time.

@@

\subsection*{Symbols}
\begin{center}
	\begin{tabular}{ l l }
$A$ & system state matrix \\
$\widetilde{A}$ & augmented system state matrix \\
$B$ & system input matrix \\
$\widetilde{B}$ & augmented system input matrix \\
$C$ & system output matrix \\
$\widetilde{C}$ & augmented system output matrix \\
$d_{\max}$ & maximal deviation from the steady state \\
TODO & odlisit max size of reference step change \\
$F$ & slope of the affine control law \\
$g$ & section of the affine control law \\
$I$ & identity matrix \\	
$k$ & step of the prediction horizon \\
$N$ & prediction horizon \\
$n_{\mathrm{u}}$ & size of system inputs \\
$n_{\mathrm{y}}$ & size of system outputs \\
$n_{\widetilde{\mathrm{x}}}$ & size of augmented system states \\
$Q_{\mathrm{x}}$ & penalty matrix of the built-in integrator \\
$Q_{\mathrm{x,L}}$ & lower bound on the penalty matrix of the built-in integrator \\
$Q_{\mathrm{x,U}}$ & upper bound on the penalty matrix of the built-in integrator \\
$Q_{\mathrm{y}}$ & penalty matrix of the control error \\
$Q_{\mathrm{y,L}}$ & lower bound on the penalty matrix of the control error \\
$Q_{\mathrm{y,U}}$ & upper bound on the penalty matrix of the control error \\
$R$ & penalty matrix of system inputs \\
$R_{\mathrm{L}}$ & lower bound on the penalty matrix of system inputs \\
$R_{\mathrm{U}}$ & upper bound on the penalty matrix of system inputs \\
$\mathcal{R}$ & critical region \\
$\mathbb{R}$ & Euclidean space of real numbers \\
$t$ & time, s \\
$t_{\epsilon}$ & settling time, s \\
$T$ & temperature, $^{\circ}\mathrm{C}$ \\
$T_{\mathrm{s}}$ & sampling time, s \\
$T^{\mathrm{s}}$ & steady state of temperature, $^{\circ}\mathrm{C}$ \\
$u$ & control inputs \\
$u_{\mathrm{L}}$ & control inputs asscociated with the lower boundary controller\\
$u_{\mathrm{U}}$ & control inputs asscociated with the upper boundary controller\\
\end{tabular}
\end{center}

@@

\begin{center}
	\begin{tabular}{ l l }
$U$ & voltage, \% \\
$U^{\mathrm{s}}$ & steady state of voltage, \% \\
$\mathcal{U}$ & set of control inputs \\
$V$ & volume of heating medium, l \\
$x$ & system states \\
$\widetilde{x}$ & augmented system states \\
$x_{\mathrm{I}}$ & system states corresponding to the built-in integrator \\
$y$ & system outputs \\
$y_\mathrm{\max}$ & maximal value of system outputs \\
$y_\mathrm{\min}$ & minimal value of system outputs \\
$y_\mathrm{ref}$ & reference value of system outputs \\	
$\mathcal{Y}$ & set of system outputs \\
\textit{0} & zero matrix
\end{tabular}
\end{center}

>>>>

In this section, the results of the proposed self-tuning method are analyzed using an experimental implementation. The self-tuning strategy utilizes tuning parameter based on the size of reference change (Section~\ref{sec:self_tunable_delta_ref}) and the scaling of tuning parameter based on splitting the interval of the parameter and assigning the interval parts to specific operating conditions (Section~\ref{sec:self_tunable_rho_scaling}).

The plant on which the control was implemented and analyzed is a laboratory counter-current liquid-to-liquid plate heat exchanger Armfield Process Plant Trainer PCT23~\cite{pct23}, see Figure~\ref{fig:HE}. The cold feed as well as heating medium are transported to the heat exchanger by two peristaltic pumps. The flow rate of the feed of the cold medium (water) is constant, while the aim of control is to track the reference value of the cold medium temperature. Therefore, the controlled variable is the feed temperature $T$ at the outlet of the heat exchanger. The associated manipulated variable is the voltage $U$ corresponding to the power of the pump feeding the heat exchanger by the hot medium (water). The voltage is within the range of [$0-5$]\,V normalized into the relative values in percentage. The manipulated variable is constrained from 20\,\% to 100\,\% to respect the physical limitations of the operating conditions. As the heat exchange is a nonlinear and asymmetric process~\cite{Liptak}, this heat exchanger represents a suitable candidate for the presented controller tuning strategy.

The matrices of the linear state-space model of the plant are
\begin{subequations}
	\label{eq:model_A_B} 
	\begin{eqnarray}
		A = \begin{bmatrix}
			0.839
		\end{bmatrix}, \quad
		B = \begin{bmatrix}
			0.039
		\end{bmatrix}, \quad
		C = \begin{bmatrix}
			1
		\end{bmatrix}, 
%		\quad
%		D = \begin{bmatrix}
%			0
%		\end{bmatrix}.
		\end{eqnarray}
\end{subequations}
considering the sampling time $T_\mathrm{s}$ = 1\,s. 
The constraints are considered in the terms of control inputs
\begin{eqnarray}
\label{eq:u_const}
	-15\% \le u \le 65\%
\end{eqnarray}
where the variable $u$ represents the control inputs in the deviation form. The values of feed temperature and voltage of the heating medium pump corresponding to zero steady state are respectively $T^\mathrm{s}$~=~35 $^{\circ}\mathrm{C}$ and $U^\mathrm{s}$~=~35\%.
To fully investigate the control performance of the pure self-tuning method, the constraints on the system states~\eqref{eq:mpc_problem_state_constraints} were omitted to prevent shaping the computed control inputs by taking into account the bounds on the controlled variable.

The penalty matrices of the problem Eq.~\eqref{eq:mpc_problem} were systematically tuned and the corresponding control was implemented on the laboratory heat exchanger for each of the considered explicit MPC setup. 
First, the aim of tuning was to determine, which penalty matrix is the most suitable for real-time tuning. Based on the set of experimentally collected data, the most significant effect on the control trajectories had tuning the penalty matrix $Q_\mathrm{y}$, while still preserving a satisfactory control performance, i.e., without steady-state control error and significant oscillations around the reference value. Next, the boundary values of the tunable matrix $Q_\mathrm{y}$ were tuned as $Q_\mathrm{y, L}$ = 100 and $Q_\mathrm{y, U}$ = 1000. The built-in integrator was penalized with the fixed penalty matrix $Q_\mathrm{x}$ = 1 and the control input with the fixed penalty matrix $R$ = 10. The prediction horizon $N$ was set to 20 control steps. The explicit MPC controllers were constructed in MATLAB R2020b using the Multi-Parametric Toolbox 3~\cite{mpt_conf}. 
The controllers were implemented to track a time-varying PWC reference.

\textcolor{red}{[ TODO: Please, consider introducing the sequence of the reference values/control scenarios here. ]}

For the initial 200 seconds, the system was in its steady-state value. After that, the reference changed its value twice upwards and twice downwards. The reference changes also acquired different sizes in order to examine the proposed tuning method as it is dependent on the size of the reference step change. The control profiles generated for the both considered boundary control setups are compared in Figure~\ref{fig:CV_boundaries} for the controlled outputs, and in Figure~\ref{fig:MV_boundaries} for the control inputs. 
Note, the constructed explicit MPC controller computed control inputs to respect the constraints on the control inputs and they need not to be saturated/truncated afterwards.

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{images/CV_boundaries}
		\caption{Controlled variable: output cold medium temperature for two boundary controllers. The solid lines represent the controlled temperature $T$ and the dashed line represents the reference value $T_{\mathrm{ref}}$.}
		\label{fig:CV_boundaries}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{images/MV_boundaries}
		\caption{Manipulated variables: input hot medium flow-rate for two boundary controllers. The solid lines represent the voltage $U$ and the dashed lines represent the constraints $U_{\mathrm{bound}}$.}
		\label{fig:MV_boundaries}
	\end{center}
\end{figure}

The trajectories in Figure~\ref{fig:CV_boundaries} show the asymmetric nature of controlling the plant of plate heat exchange mainly when observing the overshoots and undershoots. When applying the control inputs associated with the lower bound $Q_\mathrm{y, L}$ in~\eqref{eq:mpc_problem_cost}, significant undershoots are present when tracking the reference downwards, i.e., when the reference change is negative. On the contrary, when implementing the controller associated with $Q_\mathrm{y, U}$ in~\eqref{eq:mpc_problem_cost}, the undershoots are negligible, but significant overshoots can be seen when tracking the reference upwards.

These main experimental observations established the base for the strategy of self-tuning the penalty matrix $Q_\mathrm{y}$. The strategy follows the ideas summarized in Section~\ref{sec:self_tunable}. Therefore, utilizing the boundary controller with the penalty matrix $Q_\mathrm{y,L}$ is preferred when the reference changes upwards. On the contrary, the controller associated with $Q_\mathrm{y,U}$ is preferred for negative reference step changes. The splitting value of the tuning parameter was chosen in the middle of the interval, i.e., $\rho_{\mathrm{s}} = 0.5$. The remaining parameter that needed to be set was the maximal admissible size of the reference step change $d_{\max}$, which was experimentally determined to $15\,^{\circ}\mathrm{C}$.

The control results of the self-tunable technique compared to the boundary controllers can be seen in Figure~\ref{fig:CV} for the controlled outputs, and in Figure~\ref{fig:MV} for the control inputs. It can be seen that the tuned controller combined the benefits of the two boundary controllers. The overshoots and undershoots were reduced, as in the first half of control the penalty matrix $Q_\mathrm{y}$ acquired value from the first half of the penalty interval. When tracking the reference with negative step change, the penalty matrix acquired the values from the second half of the interval, i.e., closer to the upper bound $Q_\mathrm{y,U}$. 
The similarity with the boundary controllers can be seen also on the control inputs profiles. Note, the constraints on the input variable were satisfied as they were scaled using linear interpolation based on the boundary controllers which are constructed considering the input constraints.

@@

The control performance was also investigated quantitatively. The Table~\ref{tab:control_performance} summarizes the evaluated control performance criteria computed for the two boundary controllers and for the self-tuned controller. The control performance is evaluated for each reference step change separately. The considered quality criteria are: sum-of-squared based criterion -- integral squared error $ISE$, maximal overshoot/undershoot $\sigma_{\mathrm{max}}$, settling time $t_{\epsilon}$ for 5\,\%-neighbourhood of the reference temperature $T_{\mathrm{ref}}$, and the volume of the hot medium $V$ consumed for the corresponding control. To provide a better readability of the computed results in Table~\ref{tab:control_performance}, the best values, i.e., the minimum values, are emphasized using a bold font style.

\begin{table}[h!]
	\begin{center}
		\caption{Control performance criteria.}
		\label{tab:control_performance}
		\begin{tabular}{c|c|c|c|c|c} 
			Reference step change & $Q_\mathrm{y}$ & $ISE$ [$^{\circ}\mathrm{C}^2$\,s] & $\sigma_{\mathrm{max}}$\,[\%] & $t_{\epsilon}$\,[s] & $V$\,[l] \\
			\hline
			\multirow{2}{*}{ 35\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1000 & 714 & 33.50 & 16.5 & \textbf{2.12} \\
			    & 100 & 867 & 16.65 & 12.5 & 2.36 \\ 
			    & self-tuned & \textbf{678} & \textbf{15.19} & \textbf{9.5} & 2.38 \\ 
			\hline
			\multirow{2}{*}{ 45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C } & 1000 & 365 & 47.20 & \textbf{5} & \textbf{2.49} \\
			    & 100 & 606 & 23.25 & 26.5 & 3.19 \\ 
			    & self-tuned & \textbf{248} & \textbf{19.13} & 9.5 & 3.35 \\ 
			\hline
		    \multirow{2}{*}{ 50\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C } & 1000 & 245 & \textbf{18.92} & \textbf{6.5} & \textbf{2.00} \\
				& 100 & 398 & 79.64 & 31 & \textbf{2.00} \\ 
				& self-tuned & \textbf{186} & 24.59 & \textbf{6.5} & 2.10 \\ 
			\hline
			\multirow{2}{*}{ 45\,$^{\circ}$C $\rightarrow$ 35\,$^{\circ}$C } & 1000 & 1024 & 18.43 & 22.5 & 0.94 \\
				& 100 & 1402 & 41.87 & 90 & \textbf{0.93} \\ 
				& self-tuned & \textbf{967} & \textbf{16.49} & \textbf{18.5} & 1.10  
		\end{tabular}
	\end{center}
\end{table}

As can be seen in Table~\ref{tab:control_performance}, that real-time self-tuning of the explicit MPC controller helped to improve two to three criteria when tracking each reference value. The cost for this is the energy associated with control. Although the integral square error, maximal overshoot/undershoot and settling time decreased, the volume of consumed heating medium did not. Nevertheless, the average deterioration in the terms of consumed heating medium is approximately by~17\,\%.

The relative improvement in percentage of using the self-tunable controller computed subject to the second best setup is summarized in Table~\ref{tab:improvement} for each reference step change separately. The negative numbers represent deterioration compared to the best controller setup in the corresponding reference tracking.

\begin{table}[h!]
	\begin{center}
		\caption{Relative improvement of the control performance using the self-tunable explicit MPC controller.}
		\label{tab:improvement}
		\begin{tabular}{c|c|c|c|c} 
			Reference step change & $\Delta ISE$\,[\%] & $\Delta \sigma_{\mathrm{max}}$\,[\%] & $\Delta t_{\epsilon}$\,[\%] & $\Delta V$\,[\%] \\
			\hline
			35\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C &  5 &  9 & 24 &$-12$ \\ 
			% \hline
			45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C & 32 & 18 &$-90$ & $-35$ \\ 
			% \hline
			50\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C & 24 &$-30$& 0 &$-4$\\ 
			% \hline
			45\,$^{\circ}$C $\rightarrow$ 35\,$^{\circ}$C &  6 & 11 & 18 &$-18$  
		\end{tabular}
	\end{center}
\end{table}
%\begin{table}[h!]
%	\begin{center}
%		\caption{Relative improvement of the control performance using the self-tunable explicit MPC controller.}
%		\label{tab:improvement}
%		\begin{tabular}{c|c|c|c|c} 
%			Reference step change & $\Delta ISE$\,[\%] & $\Delta \sigma_{\mathrm{max}}$\,[\%] & $\Delta t_{\epsilon}$\,[\%] & $\Delta V$\,[\%] \\
%			\hline
%			35\,$^{\circ}$C $\rightarrow$ 45\,$^{\circ}$C & 5.04 & 8.77 & 24.00 & -12.26 \\ 
%			\hline
%			45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C & 32.05 & 17.72 & -90.00 & -34.54 \\ 
%			\hline
%			45\,$^{\circ}$C $\rightarrow$ 50\,$^{\circ}$C & 24.08 & -29.97 & 0 & -4.48 \\ 
%			\hline
%			45\,$^{\circ}$C $\rightarrow$ 35\,$^{\circ}$C & 5.57 & 10.53 & 17.78 & -18.28  
%		\end{tabular}
%	\end{center}
%\end{table}

Implementing a self-tunable explicit MPC controller leads to improved control performance in the most analyzed quality criteria, see Table~\ref{tab:control_performance}. In general, utilizing a proposed controller with a scalable aggressiveness according to the operating conditions leads to higher accuracy ($ISE$), lower value of the overshoots ($\sigma_{\mathrm{max}}$), and faster achieving the reference value ($t_{\epsilon}$). %It should be also noted that the deterioration of every criterium is calculated subject to the best out of \textit{two} other controllers. It follows that the evaluation is strict as in practice only one controller would be available.
Although the volume of the hot medium $V$ is not reduced in the considered control scenario, the improved control performance may lead to other benefits.

@@

\textcolor{red}{[ TODO: How about to simply omit the evaluation of the volume $V$ and write something around these lines: ]}

Compared to the considered non-self-tunable controllers, the generated control trajectories and the computed quality criteria confirmed the improved control performance for the reference tracking control problem of the heat exchanger with the non-linear and asymmetric behaviour. 
For various operating conditions, the proposed self-tunable controller outperformed the conventional controllers in the means of the squared-error-based criterion ($ISE$) of the output temperature $T$ within a range $[5-30]$\,\%.  
Together with the reduced overshoots and settling times, the achieved improvements of the control performance enables non-negligible energy savings when the precision and the operating time to ensure the product with the required properties. Although the analyzed extensive case study was realized considering the laboratory scaled heat exhcnager plant, the introduced self-tunable control policy enables significant reduction of the corresponding carbon footprint emissions in the industrial applications.

\section{Conclusion}
\label{sec:conclusion}

This paper deals with the experimental implementation of tunable approximated explicit model predictive control and provides a technique for self-tuning controller design. Based on the current reference value, the tuning parameter is scaled using linear interpolation. The previously published paper related to the self-tunable explicit MPC suggested tuning based on the distance of the reference value from the steady state. This paper presents a novel idea of tuning based on the size of reference step change. The tuning procedures aim to compensate for the nonlinear behavior of the controlled system. The tuning parameter is updated when the reference changes. It is calculated as the ratio between the size of the reference change and the maximal possible size of the reference change, which is specified before operation. Another novel contribution comes with splitting the interval of tuning parameter into two ranges, while both ranges are assigned to different operating conditions. The proposed method is implemented on a laboratory heat exchanger with nonlinear and asymmetric behavior. The non-symmetry makes the plant a suitable candidate for splitting the interval of the tuning parameter. The decision criterium is negativity or positivity of reference change. When the reference changed upwards, the control input was tuned in the first part of the interval and approached the boundary controller associated with the lower bound on the penalty matrix. On the contrary, when the reference changed downwards, the control input was tuned to approach the control input from the boundary controller with the upper bound on the penalty matrix. To investigate the control results properly, the control performance was evaluated quantitatively. Compared to control with only one controller, the self-tunable technique helped to decrease the maximal overshoots/undershoots, integral square error, and settling time.

@@

\subsection*{Symbols}
\begin{center}
	\begin{tabular}{ l l }
$A$ & system state matrix \\
$\widetilde{A}$ & augmented system state matrix \\
$B$ & system input matrix \\
$\widetilde{B}$ & augmented system input matrix \\
$C$ & system output matrix \\
$\widetilde{C}$ & augmented system output matrix \\
$d_{\max}$ & maximal deviation from the steady state \\
TODO & odlisit max size of reference step change \\
$F$ & slope of the affine control law \\
$g$ & section of the affine control law \\
$I$ & identity matrix \\	
$k$ & step of the prediction horizon \\
$N$ & prediction horizon \\
$n_{\mathrm{u}}$ & size of system inputs \\
$n_{\mathrm{y}}$ & size of system outputs \\
$n_{\widetilde{\mathrm{x}}}$ & size of augmented system states \\
$Q_{\mathrm{x}}$ & penalty matrix of the built-in integrator \\
$Q_{\mathrm{x,L}}$ & lower bound on the penalty matrix of the built-in integrator \\
$Q_{\mathrm{x,U}}$ & upper bound on the penalty matrix of the built-in integrator \\
$Q_{\mathrm{y}}$ & penalty matrix of the control error \\
$Q_{\mathrm{y,L}}$ & lower bound on the penalty matrix of the control error \\
$Q_{\mathrm{y,U}}$ & upper bound on the penalty matrix of the control error \\
$R$ & penalty matrix of system inputs \\
$R_{\mathrm{L}}$ & lower bound on the penalty matrix of system inputs \\
$R_{\mathrm{U}}$ & upper bound on the penalty matrix of system inputs \\
$\mathcal{R}$ & critical region \\
$\mathbb{R}$ & Euclidean space of real numbers \\
$t$ & time, s \\
$t_{\epsilon}$ & settling time, s \\
$T$ & temperature, $^{\circ}\mathrm{C}$ \\
$T_{\mathrm{ref}}$ & sampling time, s \\
$T_{\mathrm{s}}$ & sampling time, s \\
$T^{\mathrm{s}}$ & steady state of temperature, $^{\circ}\mathrm{C}$ \\
$u$ & control inputs \\
\end{tabular}
\end{center}

@@

\begin{center}
	\begin{tabular}{ l l }
$u_{\mathrm{L}}$ & control inputs asscociated with the lower boundary controller\\
$u_{\mathrm{U}}$ & control inputs asscociated with the upper boundary controller\\
$U$ & voltage, \% \\
$U_{\mathrm{bound}}$ & steady state of voltage, \% \\
$U^{\mathrm{s}}$ & steady state of voltage, \% \\
$\mathcal{U}$ & set of control inputs \\
$V$ & volume of heating medium, l \\
$x$ & system states \\
$\widetilde{x}$ & augmented system states \\
$x_{\mathrm{I}}$ & system states corresponding to the built-in integrator \\
$y$ & system outputs \\
$y_\mathrm{\max}$ & maximal value of system outputs \\
$y_\mathrm{\min}$ & minimal value of system outputs \\
$y_\mathrm{ref}$ & reference value of system outputs \\	
$\mathcal{Y}$ & set of system outputs \\
\textit{0} & zero matrix
\end{tabular}
\end{center}
